RDF Validation Requirements - Evaluation and Logical Underpinning

### 2015-01-06 paper/RDF validation requirements/RDF validation requirements.tex
add descriptions

\begin{DL}
$parentOf \sqsubseteq \neg parentOf^{-}$ or\\
$parentOf \sqcap parentOf^{-} \sqsubseteq \bot$ or \\
disjoint(parentOf, parentOf$^{-}$)
\end{DL}

@@

\begin{itemize}
	\item child parent relations (dbo:child) cannot be symmetric
\end{itemize}

@@

\begin{itemize}
	\item person birth place relations (dbo:birthPlace) cannot be symmetric
\end{itemize}

@@

\begin{itemize}
  \item DL terminology: local reflexivity
	\item set of individuals that are related to themselves via a given role \cite{Kroetzsch2012}.
	\item set of individuals (of a specific class) that are related to themselves via a given object property
\end{itemize}

@@

\begin{itemize}
	\item set of individuals that are talking to themselves
\end{itemize}

@@

\begin{itemize}
	\item R-59-REFLEXIVE-OBJECT-PROPERTIES
\end{itemize}

\begin{itemize}
  \item DL terminology: reflexive roles, global reflexivity
	\item global reflexivity can be expressed by imposing local reflexivity on the top concept \cite{Kroetzsch2012}
\end{itemize}

@@

@@

\begin{itemize}
	\item R-60-IRREFLEXIVE-OBJECT-PROPERTIES
\end{itemize}

\begin{itemize}
	\item DL terminology: irreflexive roles
  \item A role is irreflexive if it is never locally reflexive \cite{Kroetzsch2012}
	\item An object property irreflexivity axiom IrreflexiveObjectProperty( OPE ) states that the object property expression OPE is irreflexive — that is, no individual is connected by OPE to itself. 
\end{itemize}

@@

\begin{DL}
$\top$ $\sqsubseteq$ $\neg$ $\exists  marriedTo . Self$   OR\\
$\exists marriedTo . Self \sqsubseteq \bot$  OR\\
$marriedTo \sqcap marriedTo^{-} \sqsubseteq \bot$ (without special self-concept Self)
\end{DL}

@@

\begin{itemize}
	\item a resource cannot be its own parent (dbo:parent)
\end{itemize}

@@

\begin{itemize}
	\item a resource cannot be its own child (dbo:child)
\end{itemize}

@@

\begin{itemize}
	\item R-66-PROPERTY-GROUPS
\end{itemize}

@@

\begin{ex}
# ShEx
# ----
<Human> { 
    (  
        foaf:name xsd:string ,
        foaf:givenName xsd:string 
    ) ,
    (
        foaf:mbox IRI ,
        foaf:homepage foaf:Document
    ) }
\end{ex}

@@

\begin{itemize}
	\item R-11-CONTEXT-SPECIFIC-EXCLUSIVE-OR-OF-PROPERTIES
\end{itemize}

\begin{itemize}
  \item exclusive or: logical operation that outputs true whenever both inputs differ (one is true, the other is false).
	\item in OWL 2, inclusive OR of properties would be possible to express, but not exclusive OR of data properties
\end{itemize}

DL (general):

@@

DL (general) (alternative):

@@

DSCL (ShEx):
\begin{ex}
<Human> { (  
    foaf:name xsd:string | 
    foaf:givenName xsd:string ) }
\end{ex}

DSCL (OWL 2):
\begin{ex}
:Human owl:disjointUnionOf ( :CC1 :CC2 ) .

:CC1 rdfs:subClassOf [
    a owl:Restriction ;
    owl:onProperty foaf:name ;
    owl:someValuesFrom xsd:string ] .
:CC2 rdfs:subClassOf [
    a owl:Restriction ;
    owl:onProperty foaf:givenName ;
    owl:someValuesFrom xsd:string ] .
\end{ex}

@@

\begin{ex}
# ShEx
# ----
<FeelingForce> { (  
    :attackingBySword xsd:boolean | 
    :attackingByForce xsd:boolean ) }
\end{ex}

\begin{ex}
# OWL 2 DL
# --------
:FeelingForce owl:disjointUnionOf ( :CC1 :CC2 ) . 
:CC1 rdfs:subClassOf [
    a owl:Restriction ;
    owl:onProperty :attackingBySword ;
    owl:someValuesFrom xsd:boolean ] .
:CC2 rdfs:subClassOf [
    a owl:Restriction ;
    owl:onProperty :attackingByForce ;
    owl:someValuesFrom xsd:boolean ] .
\end{ex}

\begin{itemize}
	\item :FeelingForce $\equiv$ 1 :attackingBySword (range xsd:boolean) XOR 1 :attackingByForce (range xsd:boolean)
\end{itemize}

\begin{DL}
A $\equiv$ $\exists$ :attackingBySword . xsd:boolean \\
B $\equiv$ $\exists$ :attackingByForce . xsd:boolean \\ 
:FeelingForce $\sqsubseteq$ ($\neg$ A $\sqcap$ B) $\sqcup$ (A $\sqcap$ $\neg$ B) \\
\end{DL}

@@

DSCL (ShEx):

\begin{ex}
<Human> { (  
    foaf:name xsd:string | foaf:givenName xsd:string+ , 
    foaf:familyName xsd:string ) }
\end{ex}

@@

\begin{itemize}
	\item R-202-CONTEXT-SPECIFIC-INCLUSIVE-OR-OF-PROPERTIES
\end{itemize}

\begin{itemize}
	\item inclusive or: A logical connective joining two or more predicates that yields the logical value "true" when at least one of the predicates is true.
	\item context can be a class $\rightarrow$ constraint applies for individuals of this specific class 
\end{itemize}

>>>>

The property \ms{parentOf} is asymmetric:

\begin{DL}
$parentOf \sqsubseteq \neg parentOf^{-}$ 
\end{DL}

alternatively:

\begin{DL}
$parentOf \sqcap parentOf^{-} \sqsubseteq \bot$ 
\end{DL}

alternatively:

\begin{DL}
disjoint(parentOf, parentOf$^{-}$)
\end{DL}

@@

Child parent relations (dbo:child) cannot be symmetric.

@@

Person birth place relations (dbo:birthPlace) cannot be symmetric.

@@

Using DL terminology \ms{Class-Specific Reflexive Object Properties} is called local reflexivity - a set of individuals (of a specific class) that are related to themselves via a given role \cite{Kroetzsch2012}.

@@

Set of individuals that are talking to themselves.

@@

\ms{Reflexive Object Properties} (reflexive roles, global reflexivity in DL) corresponds to the requirement \ms{R-59-REFLEXIVE-OBJECT-PROPERTIES}.
Global reflexivity can be expressed by imposing local reflexivity on the top concept \cite{Kroetzsch2012}.

@@

Each individual knows itself:

@@

\ms{Irreflexive Object Properties} (irreflexive roles in DL) corresponds to the requirement \ms{R-60-IRREFLEXIVE-OBJECT-PROPERTIES}.
A role is irreflexive if it is never locally reflexive \cite{Kroetzsch2012}.
An object property irreflexivity axiom IrreflexiveObjectProperty( OPE ) states that the object property expression OPE is irreflexive - that is, no individual is connected by OPE to itself.

@@

\begin{DL}
$\top$ $\sqsubseteq$ $\neg$ $\exists  marriedTo . Self$ 
\end{DL}

alternatively:

\begin{DL}
$\exists marriedTo . Self \sqsubseteq \bot$  
\end{DL}

alternatively:

\begin{DL}
$marriedTo \sqcap marriedTo^{-} \sqsubseteq \bot$ (without special self-concept Self)
\end{DL}

@@

A resource cannot be its own parent (dbo:parent).

@@

A resource cannot be its own child (dbo:child).

@@

\ms{Context-Specific Property Groups} corresponds to the requirement \\
\ms{R-66-PROPERTY-GROUPS} to group data and object properties.

@@

\begin{ex}
# ShEx:
<Human> { 
    (  
        foaf:name xsd:string ,
        foaf:givenName xsd:string 
    ) ,
    (
        foaf:mbox IRI ,
        foaf:homepage foaf:Document
    ) }
\end{ex}

@@

\ms{Context-Specific Exclusive OR of Properties} correponds to the requirement
\ms{R-11-CONTEXT-SPECIFIC-EXCLUSIVE-OR-OF-PROPERTIES}.
Exclusive or is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
%In OWL 2, inclusive OR of properties would be possible to express, but not exclusive OR of data properties
This constraint is generally expressed in DL as follows:

@@

and alternatively:

@@

\begin{ex}
# ShEx:
<Human> { (  
    foaf:name xsd:string | 
    foaf:givenName xsd:string ) }
\end{ex}

\begin{ex}
# OWL 2:
Human owl:disjointUnionOf ( :CC1 :CC2 ) .

CC1 rdfs:subClassOf [
    a owl:Restriction ;
    owl:onProperty foaf:name ;
    owl:someValuesFrom xsd:string ] .
CC2 rdfs:subClassOf [
    a owl:Restriction ;
    owl:onProperty foaf:givenName ;
    owl:someValuesFrom xsd:string ] .
\end{ex}

@@

\begin{ex}
# ShEx:
<FeelingForce> { (  
    attackingBySword xsd:boolean | 
    attackingByForce xsd:boolean ) }
\end{ex}

\begin{ex}
# OWL 2 DL:
FeelingForce owl:disjointUnionOf ( CC1 CC2 ) . 
CC1 rdfs:subClassOf [
    a owl:Restriction ;
    owl:onProperty attackingBySword ;
    owl:someValuesFrom xsd:boolean ] .
CC2 rdfs:subClassOf [
    a owl:Restriction ;
    owl:onProperty attackingByForce ;
    owl:someValuesFrom xsd:boolean ] .
\end{ex}

This means that \ms{FeelingForce} individuals are individuals having one \ms{attackingBySword} relationship (range xsd:boolean) or one \ms{attackingByForce} relationship (range xsd:boolean) but not both.

\begin{DL}
A $\equiv$ $\exists$ attackingBySword.xsd:boolean \\
B $\equiv$ $\exists$ attackingByForce.xsd:boolean \\ 
FeelingForce $\sqsubseteq$ ($\neg$ A $\sqcap$ B) $\sqcup$ (A $\sqcap$ $\neg$ B) \\
\end{DL}

@@

\begin{ex}
# ShEx:
<Human> { (  
    foaf:name xsd:string | foaf:givenName xsd:string+ , 
    foaf:familyName xsd:string ) }
\end{ex}

@@

\ms{Context-Specific Inclusive OR of Properties} corresponds to the requirement
\ms{R-202-CONTEXT-SPECIFIC-INCLUSIVE-OR-OF-PROPERTIES}.
Inclusive or is a logical connective joining two or more predicates that yields the logical value "true" when at least one of the predicates is true.
The context can be a class, i.e., the constraint applies for individuals of this specific class.
